---
title: "05) Wind direction QC"
author: "Stenka Vulova"
date: "27 Oct. 2021"
output:
  pdf_document: default
  html_document: default
---

# Libraries 

```{r libraries, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(plyr)
library(dplyr)
library(grid)
library(spatstat)
library(sf)
library(readr)
library(RColorBrewer)
library(reshape2)
library(scales)

library(tictoc) # benchmarking 
library(spatialEco) # for weighted_footprint function 

library(openair)
library(water)
library(stats)
library(zoo)
library(water)

library(MeTo)

```

The older package is no longer in the CRAN. 
https://cran.r-project.org/src/contrib/Archive/FREddyPro/ 

https://support.rstudio.com/hc/en-us/articles/219949047-Installing-older-versions-of-packages 

```{r Freddypro library }

# https://cran.r-project.org/src/contrib/Archive/FREddyPro/FREddyPro_1.0.tar.gz

#packageurl <- "https://cran.r-project.org/src/contrib/Archive/FREddyPro/FREddyPro_1.0.tar.gz"
#install.packages(packageurl, repos=NULL, type="source")

library(FREddyPro)
  
```

# Load data 

```{r load data}

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211026_QC/TUCC/04_apply_QC/TUCC_rain.RData")

nrow(TUCC_rain)

boxplot(TUCC_rain$ET)

```

# Wind direction filter 

**Vulova et al. 2021 says 17 to 35. - DOUBLE CHECK!** 

Fred, 28 Oct. 2021: "at TUCC the sonic points at 206 deg - > the mounting setup is at 26 deg and we exclude +9 deg and - 9 deg ---> sector 17 degrees –35 degrees can be excluded."

I am also excluding 17 degrees and 35 degrees themselves. 

```{r wind direction filter }

TUCC_wind = subset(TUCC_rain, wind_dir < 17 | wind_dir > 35 )
nrow(TUCC_wind) # 15700 obs. 

```

# Save TUCC data 

The whole datasets refer to those where the time period or columns are not further subsetted for modelling. However, all QC is already applied. 

```{r save data }

TUCC_whole = TUCC_wind

glimpse(TUCC_whole)

#save(TUCC_whole, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211026_QC/TUCC/05_winddir_QC/TUCC_whole.RData")

#write.csv(TUCC_whole, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211026_QC/TUCC/05_winddir_QC/TUCC_whole.csv", row.names = FALSE )

```


