---
title: "LCZ (LCZ Generator)"
author: "Stenka Vulova"
date: "17 Feb. 2022"
output: html_document
---

# Libraries 

```{r libraries, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(pbapply)

library(terra)

```

# Import the raster

Gaussian-filtered version (100-m res)

"The resulting LCZ map provided to the user is based on allTAs (100% of the TA polygons) and input features. A filteredversion is also provided using the morphological Gaussian filterdescribed in more detail inDemuzere et al. (2020a). Thisis preferred over the WUDAPT’s traditional majority post-classification, as it accounts for the distance from the centerof the kernel and differences in the typical patch size betweenclasses. For example, linear features like rivers are typicallyremoved by the majority filter. " (Demuzere et al. 2021)

"The LCZ map, its Gaussian-filtered version and the probability map are provided to theuser as a single .tifwith three bands: “lcz,” “lczFilter,” and“classProbability,” respectively." (Demuzere et al. 2021)  

```{r import LCZ }

# LCZ_Gen_Fenner_filter= LCZ[[2]]

LCZ = terra::rast("D:/Stenka_UWI/Topic_3/03_RAW_DATA/LCZ_Fenner_Generator/LCZ_Gen_Fenner_filter.tif")

terra::plot(LCZ)

```

# Polygonize 

Let's do it with `terra`. 

```{r convert to polygons }

tic("as.polygons")

LCZ_poly = terra::as.polygons(x = LCZ, trunc = TRUE,
                                     dissolve = TRUE,
                                     values = TRUE,
                                     na.rm = TRUE, 
                                     extent = FALSE )

toc()

tic.clearlog()

# as.polygons: 0.88 sec elapsed

#terra::plot(LCZ_poly)

LCZ_poly

values_LCZ = terra::values(LCZ_poly)

values_LCZ

#0				
#2				
#4				
#5				
#6				
#8				
#9				
#11				
#12				
#13	
#14				
#16				
#17	

# LCZ 4 is in there. (open highrise)

# Checking Fenner et al. 2017:
# Missing should be: 1, 3, 7, 10, and E 
# Matches this well

#terra::writeVector(x = LCZ_poly, "D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20220217_LCZ_shape_Gene/LCZ_Generator_filter.shp",
#                   filetype = "ESRI Shapefile")

```

"LCZ  1  (compact  high-rise),  3  (compactlow-rise), 7 (lightweight low-rise), 10 (heavy industry),and E (bare rock or paved) do not occur (Fig.2)." (Fenner et al. 2017)