---
title: "Berlin impervious surface fraction (1-m resolution)"
author: "Stenka Vulova"
date: "23 August 2021"
output: html_document
---

I am re-doing the Berlin impervious surface fraction.


This time, I am using the **Urban Atlas 2018** LULC data. 

I am making the raster with 1-m resolution (co-registered to 1-m BH data). 

```{r libraries, include=FALSE}

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(FSA) # Dunns test
library(RColorBrewer)
library(reshape2)
library(scales)
library(tidyverse)

# machine learning libraries 
library(caret)
#library(CAST)
library(doSNOW) # do training in parallel
library(gbm)

# https://cran.rstudio.com/web/packages/fasterize/vignettes/using-fasterize.html 

library(fasterize)
library(sf)
library(tictoc)

```

# Data import and preprocessing

## Berlin extent shapefile 

```{r import Berlin shape, echo = FALSE }

berlin_shape <- readOGR(dsn='D:/Stenka_UWI/Topic_3/03_RAW_DATA/Urban_Atlas_2018_LULC/DE001L1_BERLIN_UA2018_v013/Data',layer='Berlin_boundary') # read shapefile with Berlin extent 

# change CRS of berlin_shape to match USGS products 
berlin_shape <- spTransform(berlin_shape, CRS("+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# new Berlin shape with longitude and latitude (not UTM)
#pro_berlin_shape <- spTransform(berlin_shape, CRS("+proj=longlat +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))

plot(berlin_shape)

```

## Urban Atlas 2018

https://land.copernicus.eu/local/urban-atlas/urban-atlas-2018?tab=metadata  

Spatial resolution:  
Minimum Mapping Width: 10 m  

According to the metadata, the "Projection" is "EPSG: 3035" and "ETRS89/ LAEA Europe." 

```{r import Urban Atlas data }

UrbanAtlas = readOGR(dsn='D:/Stenka_UWI/Topic_3/03_RAW_DATA/Urban_Atlas_2018_LULC/DE001L1_BERLIN_UA2018_v013/Data', layer='DE001L1_BERLIN_UA2018')

crs(UrbanAtlas)
# CRS arguments:
# +proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs 

UA_UTM <- spTransform(UrbanAtlas, CRS("+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")) 

#plot(UA_UTM)

# new shape with longitude and latitude (not UTM)

#pro_UrbanAtlas <- spTransform(UrbanAtlas, CRS("+proj=longlat +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))

```
## Impervious coefficients 

I am assigning the impervious surface cover (based on methods from my first publication).  

```{r assign surface cover values ROTH }

# import spreadsheet with surface cover values

impervious = read.csv('D:/Stenka_UWI/Topic_2/06_R_SCRIPTS/20200730_surface_cover_fraction/20191013_LULC_sealing.csv', sep=',', header=T)

# rename code column to match 2018 data
colnames(impervious)[which(colnames(impervious) == 'CODE2012')] <- 'code_2018'

```

# Rasterize impervious cover

Now let's rasterize the impervious cover for all of Berlin. I need this data to map urban ET. 

```{r merge UA and impervious }

#### crop to Berlin ####
#plot(UA_UTM)

# Crop UA LULC to Berlin shape 
UA_Berlin <- raster::crop(UA_UTM, extent(berlin_shape))  # subset to extent of Berlin shape (envelope)
#UA_Berlin <- raster::mask(UA_Berlin, berlin_shape)   # mask out all pixels that do not lie within the Berlin shape

plot(UA_Berlin)

#### merge with surface cover classes #### 

# https://stackoverflow.com/questions/35141094/attribute-values-to-polygons-in-spatialpolygonsdataframe
new_UA_Berlin = sp::merge(UA_Berlin, impervious)

# Check that merging worked 

new_UA_Berlindf = new_UA_Berlin@data
glimpse(new_UA_Berlindf)

#### Convert to sf #### 

UA_Berlin_sf = sf::st_as_sf(x = new_UA_Berlin)

```

I need a raster to rasterize to. Let's use the 1-m BH data. 

```{r Import BH data}

# BH 
BH = raster(paste('D:/Stenka_UWI/Topic_2/04_PROCESSED_DATA/BH_VH_1m/BUI_DSM_Berlin_1m_noNA.tif', sep=''))
plot(BH, col=colorRampPalette(c("lightblue1", "navyblue", "forestgreen", 'yellow1', 'darkorange','red3', 'dark red'))(255), zlim = c(0,20))

BH

```

Let's also subset building height to Berlin. 

```{r crop BH to Berlin }

BH_Berlin <- raster::crop(BH, extent(berlin_shape))
plot(BH_Berlin)

```

# Fasterize 

```{r fasterize Berlin impervious }

tic("Fasterize in Berlin to impervious surface fraction")

impervious_Berlin = fasterize::fasterize(sf = UA_Berlin_sf, raster = BH_Berlin, field = "Impervious", fun = "max")

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic(). 

# reset the tictoc log 

tic.clearlog()

#writeRaster(impervious_Berlin,"D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20210823_surface_fractions/impervious/Berlin_impervious_fraction_1m.tif", options='COMPRESS=LZW')

plot(impervious_Berlin, col=colorRampPalette(c("khaki1", 'yellow1', 'darkorange','red3', 'purple4'))(255), main= "Impervious surface fraction (Berlin)",
     maxpixels=1e8)

impervious_Berlin

```

# Create ISF raster based upon 10-m Urban Atlas BH 

## Import UA BH data 

There is still only building height from 2012. 

```{r import UA 2012 BH }

BH_10m = raster::raster("D:/Stenka_UWI/Topic_1/03_RAW_DATA/Copernicus/Building_Height_2012/DE001L1_BERLIN_UA2012_DHM/DE001L1_BERLIN_UA2012_DHM.tif")

crs(BH_10m)
# CRS arguments:
# +proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs 

BH_10m_UTM <- raster::projectRaster(BH_10m, crs = "+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0") 

plot(BH_10m_UTM)

# save the raster 

#writeRaster(BH_10m_UTM ,"D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/Urban_Atlas_BH_2012/BH_10m_UTM.tif", options='COMPRESS=LZW')

```

## Rasterize in 10m

```{r fasterize Berlin impervious 10m }

tic("Fasterize in Berlin to impervious surface fraction")

ISF_Berlin_10m = fasterize::fasterize(sf = UA_Berlin_sf, raster = BH_10m_UTM, field = "Impervious", fun = "max")

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic(). 

# reset the tictoc log 

tic.clearlog()

#writeRaster(ISF_Berlin_10m,"D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20210823_surface_fractions/impervious/Berlin_impervious_fraction_10m.tif", options='COMPRESS=LZW')

plot(ISF_Berlin_10m, col=colorRampPalette(c("khaki1", 'yellow1', 'darkorange','red3', 'purple4'))(255), main= "Impervious surface fraction (Berlin)",
     maxpixels=1e8)

ISF_Berlin_10m

```

# Create water fraction raster using 2018 UA

## Assign surface cover 

I will assign surface cover, where 1 = present and 0 = absent. 

```{r assign surface cover values }

# import spreadsheet with surface cover values

surfacecover = read.csv('D:/Stenka_UWI/Topic_2/06_R_SCRIPTS/20200730_surface_cover_fraction/20200730_surface_fraction.csv', sep=',', header=T)

surfacecover

# rename code column to match 2018 data
colnames(surfacecover)[which(colnames(surfacecover) == 'CODE2012')] <- 'code_2018'

# https://stackoverflow.com/questions/35141094/attribute-values-to-polygons-in-spatialpolygonsdataframe
UA_water = sp::merge(UA_Berlin, surfacecover)

# Check that merging worked 

UA_waterdf = UA_water@data
glimpse(UA_waterdf)

## convert to sf 
water_sf = sf::st_as_sf(x = UA_water)


```

## Rasterize at 1m resolution 

```{r fasterize ROTH water 1m }

# https://cran.rstudio.com/web/packages/fasterize/vignettes/using-fasterize.html 

tic("Fasterize to water surface fraction")

water_1m = fasterize::fasterize(sf = water_sf, raster = BH_Berlin, field = "Water", fun = "max")

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic(). 

# reset the tictoc log 

tic.clearlog()

#writeRaster(water_1m,"D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20210823_surface_fractions/water/Berlin_water_fraction_1m.tif", options='COMPRESS=LZW')

plot(water_1m, col=colorRampPalette(c("khaki1", 'yellow1', 'darkorange','red3', 'purple4'))(255), main= "Water surface fraction")

```

## Rasterize at 10m resolution 

```{r fasterize ROTH water 10m }

# https://cran.rstudio.com/web/packages/fasterize/vignettes/using-fasterize.html 

tic("Fasterize in ROTH to water surface fraction")

water_10m = fasterize::fasterize(sf = water_sf, raster = BH_10m_UTM, field = "Water", fun = "max")

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic(). 

# reset the tictoc log 

tic.clearlog()

#writeRaster(water_10m,"D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20210823_surface_fractions/water/Berlin_water_fraction_10m.tif", options='COMPRESS=LZW')

plot(water_10m, col=colorRampPalette(c("khaki1", 'yellow1', 'darkorange','red3', 'purple4'))(255), main= "Water surface fraction")

```