---
title: "Coregister predictors (2019)"
author: "Stenka Vulova"
date: "2 Nov. 2021"
output: html_document
---

I am coregistering all predictor geodata. 

Since the final maps in my paper should be in lon-lat, and my NDVI files are already in lon-lat:  
It makes more sense to save the UTC geodata as lon-lat. 

# Libraries 

```{r libraries, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(pbapply)

library(spatialEco) # for weighted_footprint function 

devtools::install_github("AlbyDR/rSCOPE")

library(rSCOPE)

```

# NDVI 

```{r load NDVI }

NDVI_2019 = raster::brick("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211004_NDVI_interpolated/2019/NDVI_Berlin_daily_2019.grd")

plot(NDVI_2019[[207]])

NDVI_2019_DOY207 = NDVI_2019[[207]]

NDVI_2019_DOY207

#writeRaster(NDVI_2019_DOY207,
#            "D:/Stenka_UWI/Topic_3/06_R_SCRIPTS/20211102_coreg_geodata/NDVI_2019_DOY207.tif",
#            options='COMPRESS=LZW')

# I checked in QGIS: the original NDVI data is indeed 10 m x 10 m resolution!

```


# Building height (BH)

(Urban Atlas) (used in Vulova et al. 2020) (10-m resolution) -> there is still only a 2012 version, so using that
https://land.copernicus.eu/local/urban-atlas 

```{r load BH }

BH = raster::raster("D:/Stenka_UWI/Topic_3/03_RAW_DATA/Building_Height_2012/DE001L1_BERLIN_UA2012_DHM/DE001L1_BERLIN_UA2012_DHM.tif")

plot(BH, col=colorRampPalette(c("lightblue1", "navyblue", "forestgreen", 'yellow1', 'darkorange','red3', 'dark red'))(255),
     zlim = c(0,25))
BH

```

## Change CRS

```{r BH CRS }

crs(BH)
crs(NDVI_2019_DOY207)

BH_longlat = raster::projectRaster(BH, crs = crs(NDVI_2019_DOY207),
                               method = "bilinear")

BH_longlat

ncell(BH_longlat)
ncell(NDVI_2019_DOY207)
# NDVI raster is bigger. 

#writeRaster(BH_longlat,
#            "D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/01_longlat/BH_longlat.tif",
 #           options='COMPRESS=LZW')

```

# Vegetation height (VH)

(used in Vulova et al. 2021) (1-m resolution) from Berlin Environmental Atlas 

```{r load VH }

VH = raster::raster("D:/Stenka_UWI/Topic_2/04_PROCESSED_DATA/BH_VH_1m/VEG_DSM_Berlin_1m_noNA.tif")

plot(VH, col=colorRampPalette(c("lightblue1", "navyblue", "forestgreen", 'yellow1', 'darkorange','red3', 'dark red'))(255), zlim = c(0,30))
VH

```
## Change CRS

```{r VH CRS }

crs(VH)
crs(NDVI_2019_DOY207)

VH_longlat = raster::projectRaster(VH, crs = crs(NDVI_2019_DOY207),
                               method = "bilinear")

VH_longlat

ncell(VH_longlat)
ncell(NDVI_2019_DOY207)
# VH raster is bigger. 

#writeRaster(VH_longlat,
#            "D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/01_longlat/VH_longlat.tif",
#            options='COMPRESS=LZW')

```

# Impervious surface fraction (ISF)

(Urban Atlas) (polygons) already re-derived using Urban Atlas 2018 

```{r load ISF }

ISF = raster::raster("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20210823_surface_fractions/impervious/Berlin_impervious_fraction_1m.tif")

plot(ISF, col=colorRampPalette(c("lightblue1", "navyblue", "forestgreen", 'yellow1', 'darkorange','red3', 'dark red'))(255))

ISF

```

## Change CRS

```{r ISF CRS }

crs(ISF)
crs(NDVI_2019_DOY207)

tic("project raster")

ISF_longlat = raster::projectRaster(ISF, crs = crs(NDVI_2019_DOY207),
                               method = "bilinear")

toc(log = TRUE)

# reset the tictoc log 
tic.clearlog()

# project raster: 2991.31 sec elapsed

ISF_longlat

ncell(ISF_longlat)
ncell(NDVI_2019_DOY207)

#writeRaster(ISF_longlat,
 #           "D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/01_longlat/ISF_longlat.tif",
 #           options='COMPRESS=LZW')

```

# Vegetation fraction (VF)

using Alby’s dataset based upon Berlin Environmental Atlas data "green volume"

```{r load VF }

VF = raster::raster("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20210823_surface_fractions/vegfrac_ADR/vegfrac_1m.tif")

plot(VF, col=colorRampPalette(c("lightblue1", "navyblue", "forestgreen", 'yellow1', 'darkorange','red3', 'dark red'))(255))
VF

```

## Change CRS

```{r VF CRS }

crs(VF)
crs(NDVI_2019_DOY207)

tic("project raster")

VF_longlat = raster::projectRaster(VF, crs = crs(NDVI_2019_DOY207),
                               method = "bilinear")

toc(log = TRUE)

# reset the tictoc log 
tic.clearlog()

# project raster: 3238.01 sec elapsed

VF_longlat

ncell(VF_longlat)
ncell(NDVI_2019_DOY207)

#writeRaster(VF_longlat,
#            "D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/01_longlat/VF_longlat.tif",
#            options='COMPRESS=LZW')

```