---
title: "ETo (2m wind speed)"
author: "Stenka Vulova"
date: "18 Oct. 2021"
output:
  pdf_document: default
  html_document: default
---

This time, I am converting the 10m wind speed to 2m.

http://www.fao.org/3/X0490E/x0490e07.htm#wind%20speed

"Wind speeds measured at different heights above the soil surface are different. Surface friction tends to slow down wind passing over it. Wind speed is slowest at the surface and increases with height. For this reason anemometers are placed at a chosen standard height, i.e., 10 m in meteorology and 2 or 3 m in agrometeorology. For the calculation of evapotranspiration, wind speed measured at 2 m above the surface is required. To adjust wind speed data obtained from instruments placed at elevations other than the standard height of 2m, a logarithmic wind speed profile may be used for measurements above a short grassed surface:" 


# Libraries 

```{r libraries, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(plyr)
library(dplyr)
library(grid)
library(spatstat)
library(sf)
library(readr)
library(RColorBrewer)
library(reshape2)
library(scales)

library(tictoc) # benchmarking 
library(spatialEco) # for weighted_footprint function 

library(openair)
library(water)
library(stats)
library(zoo)
library(water)

library(MeTo)

```

# Calculate ET0 from DWD data 

I am calculating hourly ET0 from DWD data using the water R package.    
https://rdrr.io/cran/water/man/hourlyET.html 

I am looking at their 2016 publication on the water package:   
 "The main input arguments are the csv file and a vector with the order of the needed variables (radiation, temperature, wind speed and relative humidity) called columns."    
 
"The argument ET allows to select between two ETr methods. ET= ETo, similar to clipped, cool-season grass, is the method for short crops and ET = ETr, similar to 0.5 m tall full-cover alfalfa, is the method for tall crops. By default, it will use ET = ETr, but the user should choose according to the conditions of the weather station."  

They refer to Allen et al. (2005). Here it is:    

"R. G. Allen, I. A. Walter, E. R., T. Howell, D. Itenfisu, and M. Jensen. The ASCE standardized reference evapotranspiration equation. Technical report, 2005. [p1, 5, 11]"  


# Create dataset suitable to water package 

**Radiation**  
DWD originally gives its data "FG_STRAHL" as J/cm^2 (hourly sum of solar incoming radiation). But Alby already converted them: "The radiation variables were converted from J/cm^2 to W m-2." W m^-2 is exactly what I need for solar radiation for the water package. 

You can see in Alby's code `codeDWD_TUCCdata`:  
`# convert J/cm^2 to W m-2`

Solar radiation is shortwave. "Solar energy enters our atmosphere as shortwave radiation in the form of ultraviolet (UV) rays (the ones that give us sunburn) and visible light." https://climate.ncsu.edu/edu/RadiationTypes
"Therefore, the sun gives off *shortwave radiation*, as it is extremely hot and has a lot of energy to give."  

**Air temperature**  
DWD provides 2m air temperature in degrees Celsius.  
http://ftp-cdc.dwd.de/climate_environment/CDC/observations_germany/climate/hourly/air_temperature/historical/DESCRIPTION_obsgermany_climate_hourly_tu_historical_en.pdf

The expected units are degrees C for the water package .

**Wind speed**
DWD provides mean wind speed in m/s.  
Alby used "wind_synop": http://ftp-cdc.dwd.de/climate_environment/CDC/observations_germany/climate/hourly/wind_synop/historical/ 

The expected units are m/s for wind speed for the water package.

**Relative humidity**  
DWD provides 2m relative humidity in %.  
http://ftp-cdc.dwd.de/climate_environment/CDC/observations_germany/climate/hourly/air_temperature/historical/DESCRIPTION_obsgermany_climate_hourly_tu_historical_en.pdf  

The expected units are % for RH for the water package.  

# Import DWD data 

```{r Import DWD data }

#### air temp ####

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/20210830_DWDdata/Ta.RData")

#### RH ####

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/20210830_DWDdata/rh.RData")

#### wind speed ####

# 10m data - already all converted to 10m

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/20210830_DWDdata/ws_10m.RData")

#### solar radiation Rs ####

# it is in W m-2
# Rs 

# Rs_sun19j[i] <- (0.25 + 0.5*sun_19_C[i]/60)*Ra_sun19[i]*100
# convert to w m-2
# Rs_sun19[i] <- round((Rs_sun19j[i]*100*100)/(60*60),2)

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/20210830_DWDdata/Rs_sun.RData")

```

# Convert wind speed to 2 m

http://www.fao.org/3/X0490E/x0490e07.htm#wind%20speed

"For the calculation of evapotranspiration, wind speed measured at 2 m above the surface is required."

## Melt df 

```{r melt ws df }

ws_melt = reshape::melt(ws_19_10m, id = c("timestamp"))

glimpse(ws_melt)

# Rename columns to match the requirements from the "water" package 
names(ws_melt)[names(ws_melt) == "value"] <- "wind"
names(ws_melt)[names(ws_melt) == "variable"] <- "station_ID"

```


## Convert ws to 2m

```{r convert ws to 2m }

# https://rdrr.io/cran/MeTo/man/adj_u2.html
# Adjust wind speed data to 2 meter height. 
# u = measured wind speed at uz above ground surface [m/s]
# uz = height of windspeed measurement above ground surface [m]

ws_melt$wind_2m = MeTo::adj_u2(u = ws_melt$wind, uz = 10)

```

## Cast df 

```{r cast df }

ws_melt2 = ws_melt
ws_melt2$wind = NULL

ws_2m = dcast(ws_melt2,
              timestamp ~ station_ID, value.var = "wind_2m")

# save ws at 2m 

#write.csv(ws_2m, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/ws_2m.csv", row.names = FALSE)

#save(ws_2m, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/ws_2m.RData")

```


# Load the DWD metadata 

Here is an example from just Tegel:  

Stations_id von_datum bis_datum Stationshoehe geoBreite geoLaenge Stationsname Bundesland  

00430 19860101 20200909             36     52.5644   13.3088 Berlin-Tegel                             Berlin  


From air temp. DWD metadata:  

* Stations_ID: 430
* Stationsname: Berlin-Tegel
* Geo. Laenge [Grad]: 13.31 
* Geo. Breite [Grad]: 52.56 
* Stationshoehe: 36
* Geberhoehe ueber Grund [m]: 2

I will just load metadata from air temperature.

```{r Load the DWD metadata }

# airT + RH
load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/20210830_DWDdata/metadata/station_70km_Ta.RData")

# wind
load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/20210830_DWDdata/metadata/station_70km_Wind.RData")

# sun -> solar radiation was derived from sunshine duration
load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/20210830_DWDdata/metadata/station_70km_sun.RData")

```

# Melt datasets 

## Ta

```{r Melt Ta }

Ta_melt = reshape::melt(Ta_19, id = c("timestamp"))

glimpse(Ta_melt)

# Rename columns to match the requirements from the "water" package 
names(Ta_melt)[names(Ta_melt) == "value"] <- "temp"
names(Ta_melt)[names(Ta_melt) == "variable"] <- "station_ID"

```

## RH

```{r Melt rh }

rh_melt = reshape::melt(rh_19, id = c("timestamp"))

glimpse(rh_melt)

# Rename columns to match the requirements from the "water" package 
names(rh_melt)[names(rh_melt) == "value"] <- "RH"
names(rh_melt)[names(rh_melt) == "variable"] <- "station_ID"

```

## Wind speed

```{r Melt ws }

glimpse(ws_melt2)

ws_2m_melt = ws_melt2

# Rename columns to match the requirements from the "water" package 
names(ws_2m_melt)[names(ws_2m_melt) == "wind_2m"] <- "wind"

```

## Solar radiation

```{r Melt Rs }

Rs_sun_melt = reshape::melt(Rs_sun19, id = c("timestamp"))

glimpse(Rs_sun_melt)

# Rename columns to match the requirements from the "water" package 
names(Rs_sun_melt)[names(Rs_sun_melt) == "value"] <- "radiation"
names(Rs_sun_melt)[names(Rs_sun_melt) == "variable"] <- "station_ID"

```

# Merge datasets 

Merge the melted datasets 

```{r Merge melted dfs }

# https://stackoverflow.com/questions/8091303/simultaneously-merge-multiple-data-frames-in-a-list

met_melt = 
list(Ta_melt, rh_melt, ws_2m_melt, Rs_sun_melt) %>% purrr::reduce(full_join)

# Add "hours" = time of the day in hours in 24hs format
met_melt$hours = lubridate::hour(met_melt$timestamp)

# Add "DOY" = day of year
met_melt$DOY = lubridate::yday(met_melt$timestamp)

```

## Add metadata 

Need lat, long, and elev

https://www.rdocumentation.org/packages/water/versions/0.8/topics/hourlyET 

```{r filter metadata }

metasun = station_70km_sun %>% 
  select(Stations_id, Stationshoehe, geoBreite, geoLaenge)

metawind = station_70km_Wind %>% 
  select(Stations_id, Stationshoehe, geoBreite, geoLaenge)

metaair = station_70km_Ta %>% 
  select(Stations_id, Stationshoehe, geoBreite, geoLaenge)

# merge all metadata

meta_all = 
list(metasun, metawind, metaair) %>% purrr::reduce(full_join)

glimpse(meta_all)

names(meta_all)[names(meta_all) == 'Stations_id'] <- 'station_ID'
names(meta_all)[names(meta_all) == 'Stationshoehe'] <- 'elev'
names(meta_all)[names(meta_all) == 'geoBreite'] <- 'lat'
names(meta_all)[names(meta_all) == 'geoLaenge'] <- 'long'



```

## Merge with metadata

```{r Merge with metadata }

meta_all$station_ID = as.character(meta_all$station_ID)

# Add ID_ 
meta_all$station_ID = paste0("ID_", meta_all$station_ID)

meta_all$station_ID = as.factor(meta_all$station_ID)

met_ready = dplyr::full_join(met_melt, meta_all)

```

# Run ETo

```{r Run ETo }

# "ETo" for short crops, similar to clipped, cool-season grass

# # height_Ta_ground is always 2m (as Alby noted)

met_ready$ETo = water::hourlyET(WeatherStation = met_ready,
                              hours = met_ready$hours,
                DOY = met_ready$DOY, 
                ET.instantaneous = FALSE,
                ET = "ETo", # short crops, similar to clipped, cool-season grass
                height = 2, # weather station sensors height in meters
                lat = met_ready$lat ,
                long = met_ready$long ,
                elev = met_ready$elev
                )

```

## save ETo

```{r save ETo}

ETo_2m_melt = met_ready

#write.csv(ETo_2m_melt, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/ETo_2m_melt.csv", row.names = FALSE)

#save(ETo_2m_melt, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/ETo_2m_melt.RData")

```

# Make wide ETo df

Let's un-melt the ETo df (for interpolation).

```{r cast ETo }

glimpse(met_ready)

ETo_cast = met_ready %>% 
  select(timestamp, station_ID, ETo)

summary(ETo_cast)

# ETo NAs :271915

# remove all rows with NAs

ETo_cast2 = na.omit(ETo_cast)
summary(ETo_cast2)

levels(ETo_cast2$station_ID)

# How many rows per station ID?
table(ETo_cast2$station_ID)

# Actually, 8 of the stations have 0 observations now.
# 18 levels - 8 = 10 stations remain with ETo 

ggplot(ETo_cast2,
       aes(x = timestamp, y = ETo, 
           color = station_ID))+
  geom_line()

```

## Cast 

```{r cast ETo 2 }

ETo_cast3 = dcast(ETo_cast2,
                  timestamp ~ station_ID, value.var = "ETo")

```

## save cast ETo

```{r save cast ETo }

glimpse(ETo_cast3)

ETo_for_interpol_ws2 = ETo_cast3

#write.csv(ETo_for_interpol_ws2, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/ETo_for_interpol_ws2.csv", row.names = FALSE)

#save(ETo_for_interpol_ws2, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/ETo_for_interpol_ws2.RData")



```



