---
title: "Extract ETo from kriged data"
author: "Stenka Vulova"
date: "19 Oct. 2021"
output:
  pdf_document: default
  html_document: default
---

Now, I will extract ETo from the kriged dataset. 

```{r Libraries, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)

# ADR libraries

library(ggspatial)
library(gstat)
library(mapview)
library(rdwd)
library(leaflet)
library(tidyverse)
library(lubridate)
library(sf)
library(WindVerification) 
library(xts)
library(MeTo)
#install.packages('bit64')
library('bit64')
library(devtools)

# Benchmarking 
library(tictoc)

```

# Load data

## Berlin extent shapefile 

```{r import Berlin shape }

# Copernicus Berlin shape 

berlin_shape <- readOGR(dsn='/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_1/03_RAW_DATA/Copernicus/Urban_Atlas_2012/DE001L1_BERLIN/Shapefiles',layer='DE001L1_BERLIN_CityBoundary') # shapefile with Berlin extent 

crs(berlin_shape)

# change CRS of berlin_shape to match USGS products 
berlin_shape <- spTransform(berlin_shape, CRS("+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))

plot(berlin_shape)

# new Berlin shape with longitude and latitude (not UTM)
#pro_berlin_shape <- spTransform(berlin_shape, CRS("+proj=longlat +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))

#plot(pro_berlin_shape)

```

## Load kriged ETo

```{r Load kriged ETo }

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211019_halfh_kriged_ETo/krg_var3.RData")

```

## Load original data 

This was the ETo data that was kriged. 

```{r load original data }

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/06_R_SCRIPTS/20210716_ETo_interpol/station_70km_ETo.RData")
load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/Halfhourly/ETo_halfhour_ws2.RData")
```

# Create tower points

## ROTH 

```{r create ROTH }

# Site TU Berlin Rothenburg (ROTH) - site code in our database: rothenburg
#- active since 06/2018
#- (Y) Lat.: 52.457232째
#- (X) Lon.: 13.315827째

ROTH_pt = data.frame(x = 13.315827, y = 52.457232)

coordinates(ROTH_pt) = c("x", "y") 
proj4string(ROTH_pt) <- CRS("+proj=longlat +datum=WGS84")
ROTH_pt_utm <- spTransform(ROTH_pt, "+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs")

plot(berlin_shape)
plot(ROTH_pt_utm, add = TRUE)

```

## TUCC 

```{r create TUCC }

#Site TU Berlin Campus Charlottenburg (TUCC) - site code in our database: tumainroof
#- active since: 07/2014
#- Lat.: 52.512283째
#- Lon.: 13.327855째


TUCC_pt = data.frame(x = 13.327855, y = 52.512283 )

coordinates(TUCC_pt) = c("x", "y") 
proj4string(TUCC_pt) <- CRS("+proj=longlat +datum=WGS84")
TUCC_pt_utm <- spTransform(TUCC_pt, "+proj=utm +zone=33 +datum=WGS84 +units=m +no_defs")

plot(berlin_shape)
plot(TUCC_pt_utm, add = TRUE)

```

# Extract ETo 

## Check CRS is the same

```{r CRS check }

plot(krg_var3[[666]])
plot(berlin_shape, add = TRUE)

crs( krg_var3[[666]] ) 

crs(ROTH_pt_utm) 

#CRS arguments:
# +proj=utm +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0 
#CRS arguments:
# +proj=utm +zone=33 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0 


```

## ROTH

```{r extract ROTH }

n = nrow(ETo_halfhour_ws2)
n

Inputs_ROTH <- NULL

tic("Extracting ETo") # tic - Starts the timer and stores the start time and the message on the stack.

for (i in 1:n) {
  Inputs_ROTH$ETo[i] <- raster::extract(krg_var3[[i]], ROTH_pt_utm)
}

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic().
# When quiet is FALSE, prints the associated message and the elapsed time.

# Transforming SpatialPoints to the CRS of the Raster
# error message gone since I used ROTH_pt_utm! 

# Extracting ETo: 7.99 sec elapsed

plot(Inputs_ROTH$ETo)
summary(Inputs_ROTH$ETo)

```

```{r create df ROTH }

Extracted_ETo_ROTH = ETo_halfhour_ws2 %>% 
  dplyr::select(timestamp)

glimpse(Extracted_ETo_ROTH)

Extracted_ETo_ROTH$ETo_krg = Inputs_ROTH$ETo

Extracted_ETo_ROTH$Tower = "ROTH"

```

## TUCC

```{r extract TUCC }

Inputs_TUCC <- NULL

tic("Extracting ETo") # tic - Starts the timer and stores the start time and the message on the stack.

for (i in 1:n) {
  Inputs_TUCC$ETo[i] <- raster::extract(krg_var3[[i]], TUCC_pt_utm)
}

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic().
# When quiet is FALSE, prints the associated message and the elapsed time.

# Transforming SpatialPoints to the CRS of the Raster
# error message gone since I used TUCC_pt_utm! 

# Extracting ETo: 6.81 sec elapsed

plot(Inputs_TUCC$ETo)
summary(Inputs_TUCC$ETo)

```

```{r create df TUCC }

Extracted_ETo_TUCC = ETo_halfhour_ws2 %>% 
  dplyr::select(timestamp)

glimpse(Extracted_ETo_TUCC)

Extracted_ETo_TUCC$ETo_krg = Inputs_TUCC$ETo

Extracted_ETo_TUCC$Tower = "TUCC"

```


# Compare kriged ETo to DWD ETo

## ROTH

I am comparing kriged ETo extracted at ROTH to ETo calculated for Berlin-Dahlem (station ID # 403).

```{r compare ROTH }

Extracted_ETo_ROTH$ETo_Dahlem = ETo_halfhour_ws2$ID_403

cor.test(Extracted_ETo_ROTH$ETo_krg, Extracted_ETo_ROTH$ETo_Dahlem)

ggplot()+
  geom_line(data = Extracted_ETo_ROTH, aes (x = timestamp, y = ETo_Dahlem), color = "blue")+
  geom_line(data = Extracted_ETo_ROTH, aes (x = timestamp, y = ETo_krg), color = "red")

ggplot(data = Extracted_ETo_ROTH, aes(x = ETo_Dahlem, y = ETo_krg) )+
  geom_point(color = "green")+
  geom_smooth()


```

# Compare kriged ETo to DWD ETo

## TUCC

I am comparing kriged ETo extracted at TUCC to ETo calculated for Berlin-Tegel (station ID # 430).

```{r compare TUCC }

Extracted_ETo_TUCC$ETo_Tegel = ETo_halfhour_ws2$ID_430

cor.test(Extracted_ETo_TUCC$ETo_krg, Extracted_ETo_TUCC$ETo_Tegel)

ggplot()+
  geom_line(data = Extracted_ETo_TUCC, aes (x = timestamp, y = ETo_Tegel), color = "blue")+
  geom_line(data = Extracted_ETo_TUCC, aes (x = timestamp, y = ETo_krg), color = "red")

ggplot(data = Extracted_ETo_TUCC, aes(x = ETo_Tegel, y = ETo_krg) )+
  geom_point(color = "green")+
  geom_smooth()


```

# Compare 2 kriged ETos

```{r compare 2 kriged ETos }

compare_krg = Extracted_ETo_ROTH %>% 
  dplyr::select(timestamp, ETo_krg)

names(compare_krg)[names(compare_krg) == 'ETo_krg'] <- 'ETo_ROTH'
compare_krg$ETo_TUCC = Extracted_ETo_TUCC$ETo_krg

cor.test(compare_krg$ETo_TUCC, compare_krg$ETo_ROTH)

ggplot()+
  geom_line(data = compare_krg, aes (x = timestamp, y = ETo_TUCC), color = "blue")+
  geom_line(data = compare_krg, aes (x = timestamp, y = ETo_ROTH), color = "red")

ggplot(data = compare_krg, aes(x = ETo_ROTH, y = ETo_TUCC) )+
  geom_point(color = "green")+
  geom_smooth()

```

# Save kriged ETo 

```{r save kriged ETo }

Extracted_ETo_ROTH$ETo_Dahlem = NULL
Extracted_ETo_TUCC$ETo_Tegel = NULL

#save(Extracted_ETo_ROTH, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211019_halfh_kriged_ETo/Towers_extracted/Extracted_ETo_ROTH.RData")

#write.csv(Extracted_ETo_ROTH, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211019_halfh_kriged_ETo/Towers_extracted/Extracted_ETo_ROTH.csv",
    #      row.names = FALSE )

#save(Extracted_ETo_TUCC, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211019_halfh_kriged_ETo/Towers_extracted/Extracted_ETo_TUCC.RData")

#write.csv(Extracted_ETo_TUCC, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211019_halfh_kriged_ETo/Towers_extracted/Extracted_ETo_TUCC.csv",
     #     row.names = FALSE )


```

