---
title: "Mapping ET (TUCC)"
author: "Stenka Vulova"
date: "22 Nov. 2021"
output: html_document
---

Now I will map ET around the TUCC tower (1 km in each direction around the tower, or 2 km^2).  
I will use this later to validate with footprints. 

# Libraries 

```{r libraries, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(pbapply)

library(spatialEco) # for weighted_footprint function 

library(caret)
library(doSNOW) # do training in parallel
library(doParallel) 
library(randomForest)

devtools::install_github("AlbyDR/rSCOPE")

library(rSCOPE)

library(terra)

```

# NDVI 

This is NDVI cropped and resampled to the footprint. 

```{r load NDVI }

NDVI_2020 = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211004_NDVI_interpolated/2020/TUCC/NDVI_TUCC_ftp_proj_2020.tif") # UTM

NDVI_2020

plot(NDVI_2020[[188]])

```
# Other geodata

This geodata was already resampled to the footprint. 

```{r Load resampled geodata }

# VF
VF = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211021_geodata_ftp_resamp/TUCC/VF_TUCC_ftp.tif")
plot(VF)

# VH
VH = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211021_geodata_ftp_resamp/TUCC/VH_TUCC_ftp.tif")
plot(VH)

# BH
BH = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211021_geodata_ftp_resamp/TUCC/BH_TUCC_ftp.tif")
plot(BH)

# ISF
ISF = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211021_geodata_ftp_resamp/TUCC/ISF_TUCC_ftp.tif")
plot(ISF)

```


# Kriged ETo

## Original ETo data

This is the half-hourly ETo data before it was kriged.

```{r original ETo }

load("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211018_ETo/Halfhourly/ETo_halfhour_ws2.RData")

glimpse(ETo_halfhour_ws2) # 35,088 obs. 

# Which row is 26 July 2020, 14 UTC?


which(ETo_halfhour_ws2$timestamp == as.POSIXct("2020-07-26 14:00:00", tz = "UTC"))
# 9917
# 2020-07-26 14:00:00

```

I am loading the kriged ETo as UTM.

```{r Kriged ETo }

ETo = readRDS(file = "D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211019_halfh_kriged_ETo/krg_var3.rds")

length(ETo)

plot(ETo[[9917]])
ETo[[1]]

```

## Convert all ETo layers to terra format 

```{r convert all ETo layers to terra format }

# testETo = ETo[[1]]
# ETo_terra = terra::rast(testETo)

ETo_l_terra = pbapply::pblapply(
  1:length(ETo), function(i)
    terra::rast(ETo[[i]])
)

# |++++++++++++++++++++++++++++++++++++++++++++++++++| 100% elapsed=03m 13s

```
Saving the Spatraster list did not work! Just remake it every time. 

# Terra resample

`terra::resample` IS faster than the `terra::extract` approach (tested already).

```{r terra resample }

ETo_terra = terra::rast(ETo[[9917]])


tic("resample ETo to NDVI")

ETo_10m = terra::resample(x = ETo_terra,
                          y = NDVI_2020,
                          method = "bilinear")

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic().

tic.clearlog()

# resample ETo to NDVI: 0.06 sec elapsed

plot(ETo_terra)
plot(ETo_10m)
ETo_10m


```
# Import RF model 

I am predicting in 2020. 
Therefore, the model should be trained in **2019**. 

```{r import RF model }

mod = readRDS("D:/Stenka_UWI/Topic_3/05_RESULTS/20211101_RF_models/Train_2019/Mod_Train2019_RF.rds")

plot(caret::varImp(mod))

```
# Make stack

All predictors are here in footprint resolution. 

```{r create raster stack }

NDVI_20200726 = NDVI_2020[[207]]

pred_stack = c(BH, ISF, VH, VF, NDVI_20200726, ETo_10m)

names(pred_stack) = c("BH", "ISF", "VH", "VF", "NDVI_FP", "ETo_krg")

pdf("D:/Stenka_UWI/Topic_3/06_R_SCRIPTS/20211122_map_TUCC_20/pred_stack.pdf", height = 8, width = 12 )
plot(pred_stack)

dev.off()

pred_stack

```

# Df predict 

# Turn raster into df

## Terra
First with terra:

```{r turn into df terra }

tic("SpatRaster to df")

pred_stack_df = terra::as.data.frame(x = pred_stack, xy = TRUE)

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic().

tic.clearlog()

# SpatRaster to df: 0.29 sec elapsed

glimpse(pred_stack_df) # Rows: 37,973

```

```{r terra df to raster }



test_df_to_spat = terra::rast(pred_stack_df, type = "xyz")

pdf("D:/Stenka_UWI/Topic_3/06_R_SCRIPTS/20211122_map_TUCC_20/pred_stack.pdf", height = 8, width = 12,
     pointsize = 20)

plot(test_df_to_spat)

dev.off()

```


# Predict on df 

```{r predict on df }

 # df with NAs removed
pred_df = pred_stack_df
# remove columns non-predictors 
pred_df$x = NULL
pred_df$y = NULL


tic("predict on df")

pred_ET = stats::predict(object = mod, pred_df)

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic().

tic.clearlog()


# predict on df: 0.91 sec elapsed

pred_stack_df$pred_ET = pred_ET

```
# Function with date 

```{r function with date }

# CAUTION: run for 1 year at a time! (Since my NDVI rasterstack are daily rasters for 1 year).

# assumption: all geodata is already in longlat
# assumption: all geodata (except for ETo) is already coregistered 

# This function returns a dataframe (xyz) which can easily be converted to a raster later 

# NDVI_stack, VH, BH, VF, ISF: Should be coregistered SpatRasters (10-m resolution)

# NDVI_stack should be a "rasterstack" for a single year (one NDVI layer/day)

# ETo_stack: ETo SpatRaster at 1-km resolution; a stack of half-hourly ETo rasters 

# mod: Trained model 

# date: "2020-01-01 00:00:00 UTC" for example, should be in UTC

# ETo_df: Dataframe with original ETo prepared for kriging but not yet kriged "ETo_halfhour_ws2" 
# ETo_df should have the same number of rows and same timestamp as the list of kriged ETo rasters. 
# For example ETo_l_terra has 35,088 elements and ETo_halfhour_ws2 has 35088 obs. 
# ETo_l_terra is kriged version of ETo_halfhour_ws2

mapET_date = function(date, ETo_df, NDVI_stack, VH, BH, VF, ISF, ETo_stack, mod) {
  
  # select ETo based on date 
  row_ETo = which(ETo_df$timestamp == as.POSIXct(date, tz = "UTC")) # row of this date
  ETo = ETo_stack[[row_ETo]] # select ETo raster 
  
  # extract DOY for NDVI 
  DOY = lubridate::yday(date) # day of year
  NDVI = NDVI_stack[[DOY]] # select NDVI based on DOY 
  
  ETo_10m = terra::resample(x = ETo ,y = NDVI, method = "bilinear") # resample ETo to NDVI 
  
  pred_stack = c(BH, ISF, VH, VF, NDVI, ETo_10m) # create prediction raster stack
  names(pred_stack) = c("BH", "ISF", "VH", "VF", "NDVI_FP", "ETo_krg") # rename stack
  
  pred_stack_mask = terra::mask(x = pred_stack, mask = NDVI ) # Mask out NAs (outside of Berlin)
  
  pred_stack_df = terra::as.data.frame(x = pred_stack_mask, xy = TRUE) # convert to dataframe 
  
  pred_df = pred_stack_df
# remove columns that are not predictors 
  pred_df$x = NULL
  pred_df$y = NULL
  
  pred_ET = stats::predict(object = mod, pred_df)
  
  pred_stack_df$pred_ET = pred_ET # add predicted ET to df with coordinates
  
  pred_df_final = pred_stack_df %>% dplyr::select(x,y, pred_ET) # coords and predicted ET
  
  return(pred_df_final)
  
}

```

# Test mapping function 

Here I am testing the function for a single timestamp.

```{r test function mapET_date }

tic("map ET")

test_function2 = mapET_date(date = as.POSIXct("2020-07-26 14:00:00", tz = "UTC"),
                       ETo_df = ETo_halfhour_ws2,
                       NDVI_stack = NDVI_2020,
                       VH = VH, BH = BH, VF = VF, ISF = ISF,
                       ETo_stack = ETo_l_terra,
                       mod = mod)

toc(log = TRUE) #Notes the current timer and computes elapsed time since the matching call to tic().

tic.clearlog()

# map ET: 1.62 sec elapsed

test_map = terra::rast(test_function2, type = "xyz")

plot(test_map)

```

# Run function with lapply

```{r Run lapply function }

# 17520 is the last timestamp in 2019. 

# 17521 is the first timestamp in 2020.


map_list = pbapply::pblapply(17521:35088 , function(i)
  mapET_date(
    date = ETo_halfhour_ws2$timestamp[i],
    ETo_df = ETo_halfhour_ws2,
                       NDVI_stack = NDVI_2020,
                       VH = VH, BH = BH, VF = VF, ISF = ISF,
                       ETo_stack = ETo_l_terra,
                       mod = mod
  )
  )

#  |++++++++++++++++++++++++++++++++++++++++++++++++++| 100% elapsed=07h 15m 06s

# save as rds file

saveRDS(map_list, "D:/Stenka_UWI/Topic_3/05_RESULTS/20211117_ET_maps_towers/2020/TUCC/ETmaps_TUCC_2020.rds")

map_list2 = readRDS("D:/Stenka_UWI/Topic_3/05_RESULTS/20211117_ET_maps_towers/2020/TUCC/ETmaps_TUCC_2020.rds")

```

I am testing the conversion to a raster. 

```{r test convert to raster }

tic("XYZ to raster")

raster_ex = raster::rasterFromXYZ(map_list[[8000]])

toc()

tic.clearlog()

# XYZ to raster: 2.22 sec elapsed

plot(raster_ex)

NDVI_normal = raster::raster(NDVI_20200726)

stack_test = raster::stack(raster_ex, NDVI_normal) # great, stacking works 

plot(stack_test)

```


