---
title: "Mapping ET (city-wide)"
author: "Stenka Vulova"
date: "24 Jan. 2022"
output: html_document
---

# Libraries 

```{r libraries, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(pbapply)

library(spatialEco) # for weighted_footprint function 

library(caret)
library(doSNOW) # do training in parallel
library(doParallel) 
library(randomForest)

#devtools::install_github("AlbyDR/rSCOPE")

#library(rSCOPE)

library(SpaDES.tools) # splitRaster

library(terra)

```

# NDVI 

```{r load NDVI }

NDVI_2019 = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211004_NDVI_interpolated/2019/NDVI_Berlin_daily_2019.grd")

NDVI_2019

terra::plot(NDVI_2019[[177]])

NDVI_2019[[177]]

```
# Other geodata

This geodata was already resampled to NDVI. 

```{r Load resampled geodata }

# VF
VF = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/02_resample/VF_proj.tif")
plot(VF)

# VH
VH = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/02_resample/VH_proj.tif")
plot(VH)

# BH
BH = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/02_resample/BH_proj.tif")
plot(BH)

# ISF
ISF = terra::rast("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211101_geodata_resamp/02_resample/ISF_proj.tif")
plot(ISF)

```


# Kriged ETo

## Original ETo data

This is the hourly ETo data before it was kriged. 

```{r original ETo }

load("D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211125_kriged_ETo_hourly/ETo_for_interpol_hourly.RData")

glimpse(ETo_for_interpol_hourly) # 17,544 obs. 



which(ETo_for_interpol_hourly$timestamp == as.POSIXct("2019-05-01 00:00:00", tz = "UTC"))

which(ETo_for_interpol_hourly$timestamp == as.POSIXct("2019-05-06 23:00:00", tz = "UTC"))

```
## Kriged ETo 

I am loading the **hourly** kriged ETo already as longlat. 

```{r Kriged ETo }

ETo = readRDS(file = "D:/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211125_kriged_ETo_hourly/ETo_l_longlat_hourly.rds")

length(ETo) # 17544

raster::plot(ETo[[1001]])

ETo[[846]]

```

## Convert all ETo layers to terra format 

```{r convert all ETo layers to terra format }

# testETo = ETo[[1]]
# ETo_terra = terra::rast(testETo)

ETo_l_terra = pbapply::pblapply(
  1:length(ETo), function(i)
    terra::rast(ETo[[i]])
)

#|++++++++++++++++++++++++++++++++++++++++++++++++++| 100% elapsed=01m 17s

```
Saving the Spatraster list did not work! Just remake it every time. 

# Import RF model 

I am predicting in 2019. 
Therefore, the model should be trained in **2020**. 

```{r import RF model }

mod = readRDS("D:/Stenka_UWI/Topic_3/05_RESULTS/20211101_RF_models/Train_2020/Mod_Train2020_RF.rds")

plot(caret::varImp(mod))

```
# Function 

This function will take as an argument a list of dates, e.g. all the dates in `ETo_for_interpol_hourly` (the original ETo dataframe before kriging). 
From each date, the function will extract a DOY to subset the NDVI raster. 
The row of the `date` selected will also be used to subset to kriged ETo. 

Only one year can be run at a time.  

## Function with date 

```{r function with date }

# assumption: all geodata is already in longlat
# assumption: all geodata (except for ETo) is already coregistered 

# This function returns a dataframe (xyz) which can easily be converted to a raster later 

# NDVI_stack, VH, BH, VF, ISF: Should be coregistered SpatRasters (10-m resolution)

# NDVI_stack should be a "rasterstack" for a single year (one NDVI layer/day)

# ETo_stack: ETo SpatRaster at 1-km resolution; a stack of half-hourly ETo rasters 

# mod: Trained model 

# date: "2019-01-01 00:00:00 UTC" for example 

# ETo_df: Dataframe with original ETo prepared for kriging but not yet kriged "ETo_for_interpol_hourly" 
# ETo_df should have the same number of rows and same timestamp as the list of kriged ETo rasters. 
# For example ETo_l_terra has 35,088 elements and ETo_for_interpol_hourly has 35088 obs. 
# ETo_l_terra is kriged version of ETo_for_interpol_hourly

mapET_date = function(date, ETo_df, NDVI_stack, VH, BH, VF, ISF, ETo_stack, mod) {
  
  # select ETo based on date 
  row_ETo = which(ETo_df$timestamp == as.POSIXct(date, tz = "UTC")) # row of this date
  ETo = ETo_stack[[row_ETo]] # select ETo raster 
  
  # extract DOY for NDVI 
  NDVI_row = lubridate::yday(date) # DOY
  NDVI = NDVI_stack[[NDVI_row]] # select NDVI based on DOY 
  
  ETo_10m = terra::resample(x = ETo ,y = NDVI, method = "bilinear") # resample ETo to NDVI (10m)
  
  pred_stack = c(BH, ISF, VH, VF, NDVI, ETo_10m) # create prediction raster stack
  names(pred_stack) = c("BH", "ISF", "VH", "VF", "NDVI_FP", "ETo_krg") # rename stack
  
  pred_stack_mask = terra::mask(x = pred_stack, mask = NDVI ) # Mask out NAs (outside of Berlin)
  
  pred_stack_df = terra::as.data.frame(x = pred_stack_mask, xy = TRUE) # convert to dataframe 
  
  pred_df = pred_stack_df
# remove columns that are not predictors 
  pred_df$x = NULL
  pred_df$y = NULL
  
  pred_ET = stats::predict(object = mod, pred_df)
  
  pred_stack_df$pred_ET = pred_ET # add predicted ET to df with coordinates
  
  pred_df_final = pred_stack_df %>% dplyr::select(x,y, pred_ET) # coords and predicted ET
  
  rast_ET = terra::rast(pred_df_final, type = "xyz")
  
  return(rast_ET)
  
}

```


# Run function with lapply 

Now I will map a week.

```{r min and max dates }

start = which(ETo_for_interpol_hourly$timestamp == as.POSIXct("2019-05-01 00:00:00", tz = "UTC"))
start

end = which(ETo_for_interpol_hourly$timestamp == as.POSIXct("2019-05-06 23:00:00", tz = "UTC"))
end 

end - start

ETo_for_interpol_hourly$timestamp[start]
ETo_for_interpol_hourly$timestamp[end]

```

```{r Run lapply function }

map_list = pbapply::pblapply(start : end, function(i)
  mapET_date(
    date = ETo_for_interpol_hourly$timestamp[i],
    ETo_df = ETo_for_interpol_hourly,
                       NDVI_stack = NDVI_2019,
                       VH = VH, BH = BH, VF = VF, ISF = ISF,
                       ETo_stack = ETo_l_terra,
                       mod = mod
  )
  )

#      |++++++++++++++++++++++++++++++++++++++++++++++++++| 100% elapsed=12h 27m 21s

# map_list List of 144


```


# Sum ET daily

This is a function that saves rasters of daily ET (mm/ day).

```{r function save_dailyET }

# Original list of predictions ("map_list") is hourly! 

# date_started_mapping = "2019-05-01"
# Date from which I ran mapping:
# on what day did the original map list start (always at 0 UTC)

# date_daily_ET = "2019-05-01"
# Date for which I want to save a daily ET sum raster 

# ETo_df = ETo_for_interpol_hourly
# ETo_df: Dataframe with original ETo prepared for kriging but not yet kriged "ETo_for_interpol_hourly" 
# ETo_df should have the same number of rows and same timestamp as the list of kriged ETo rasters. 
# For example ETo_l_terra has 17,544 elements and ETo_for_interpol_hourly has 17,544 obs. 
# ETo_l_terra is kriged version of ETo_for_interpol_hourly
# I use this dataframe to identify where a specific date is within the list of hourly ET maps. 

# map_list = map_list 
# A list of hourly ET maps (SpatRasters)

save_dailyET = function(date_started_mapping, date_daily_ET, ETo_df,
                        map_list) {
# Offset based on the date I started running mapping 
  offset = which(ETo_df$timestamp == as.POSIXct(paste(date_started_mapping, " 00:00:00", sep = ""), tz = "UTC")) - 1
  
  # Calculate where to subset 
  # First hour of the day at 0 UTC 
  start = which(ETo_df$timestamp == as.POSIXct(paste(date_daily_ET, " 00:00:00", sep = ""), tz = "UTC")) - offset
  
  # Last hour of the day at 23 UTC
  end = which(ETo_df$timestamp == as.POSIXct(paste(date_daily_ET, " 23:00:00", sep = ""), tz = "UTC")) - offset 
  
  # Subset the map list 
  map_l_date = map_list[start: end]
  
  # Make a stack 
  stack = terra::rast(map_l_date)
  
  # Sum to daily ET 
  ETdaily = terra::app(stack, fun = sum)
  
  # convert to normal raster and add CRS 
  ET_daily_r = raster::raster(ETdaily)
  raster::crs(ET_daily_r) = "+proj=longlat +datum=WGS84 +no_defs"
  
  # write the raster 
  raster::writeRaster(ET_daily_r, 
                      paste("D:/Stenka_UWI/Topic_3/05_RESULTS/20211129_dailyET_2019/June/ET_", date_daily_ET, ".tif", sep = ""),
                      options='COMPRESS=LZW' )
}

```


# Run function with list of dates 

`start = which(ETo_for_interpol_hourly$timestamp == as.POSIXct("2019-05-01 00:00:00", tz = "UTC"))`  

`end = which(ETo_for_interpol_hourly$timestamp == as.POSIXct("2019-05-06 23:00:00", tz = "UTC")) `  

```{r list of dates }

list_dates = list("2019-05-01","2019-05-02","2019-05-03","2019-05-04", "2019-05-05","2019-05-06") 

list_dates[[2]]

length(list_dates)

pbapply::pblapply(1: length(list_dates), function(i)
  save_dailyET(date_started_mapping = "2019-05-01",
             date_daily_ET = list_dates[[i]],
             ETo_df = ETo_for_interpol_hourly,
             map_list = map_list )
  )

# |++++++++++++++++++++++++++++++++++++++++++++++++++| 100% elapsed=07m 15s


```



