---
title: "Preprocessing before footprint analysis (TUCC)"
author: "Stenka Vulova"
date: "14 Oct. 2021"
output:
  pdf_document: default
  html_document: default
---

Credit is given to Justus Quanz for this code. 
https://www.hcu-hamburg.de/master/reap-resource-efficiency/academic-staff/wolfgang-dickhaut/justus-quanz 

# Libraries 

```{r libraries }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(grid)
library(spatstat)
library(sf)
library(readr)
library(RColorBrewer)
library(reshape2)
library(scales)

```

The older package is no longer in the CRAN. 
https://cran.r-project.org/src/contrib/Archive/FREddyPro/ 

https://support.rstudio.com/hc/en-us/articles/219949047-Installing-older-versions-of-packages 

```{r Freddypro library }

# https://cran.r-project.org/src/contrib/Archive/FREddyPro/FREddyPro_1.0.tar.gz

#packageurl <- "https://cran.r-project.org/src/contrib/Archive/FREddyPro/FREddyPro_1.0.tar.gz"
#install.packages(packageurl, repos=NULL, type="source")

library(FREddyPro)

```

# Correct for seasonality

## Import UMEP data  

I am importing the new UMEP point data that I created on 20 July (search distance was 1000-m).  

**Step 1)** Calculate morphometric parameters around both eddy flux towers using the QGIS plug-in UMEP for one-degree segments within a 1000-m radius around the tower.  

* Search distance (m): 1000
* Wind direction search interval (degrees): 1 
* Roughness calculation method: Kanda et al. (2013)

```{r import UMEP data }

Lambda_B_1deg <- read.csv("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_2/04_PROCESSED_DATA/20200720_UMEP_point/TUCC/TUCC_BH_1000m__IMPPoint_anisotropic.csv", header = TRUE, sep=",")

Lambda_V_1deg <- read.csv("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_2/04_PROCESSED_DATA/20200720_UMEP_point/TUCC/TUCC_VH_1000m__IMPPoint_anisotropic.csv", header = TRUE, sep=",")

Lambda_comb_1deg <- read.csv("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_2/04_PROCESSED_DATA/20200720_UMEP_point/TUCC/TUCC_comb_1000m__IMPPoint_anisotropic.csv", header = TRUE, sep=",")

glimpse(Lambda_comb_1deg)

```

## Correct UMEP data for seasonality 

Step 2) Correct zd for seasonality (Kent et al. 2017b).  

This procedure is shown in Justus’s script `RUN_Kanda_calc.R`.  

A few of Justus's functions need to be loaded beforehand. 

### planar_area_index function 

```{r planar_area_index function, include = FALSE }

# make the Kent 2017b method including vegetaion to plane area index
# @title Calculate the plane area index based on the method from Kent 2017b
# @description The function calculate the plant area index based on the plan area index
# for buildings and for vegetation (each evaluated spereatly by UMEP). The vegetation is then multipied by (1-P3D), where
# P3D is the porosity (for leaf-on  P3D = 0.2, leaf-off P3D = 0.6 and intermediate P3D = 0.4)
# @param lambda_building Give the plan area index for buildings
# @param lambda_vegetaion Give the plan area index for vegetatino
# @param P3D Give the porosity (for leaf-on  P3D = 0.2, leaf-off P3D = 0.6 and intermediate P3D = 0.4)
# @return The function returns a new plan area index including vegetation and buildings.
# @seealso Kent, C. W.; Grimmond, S.; Gatey, D. Aerodynamic roughness parameters in
#  cities: Inclusion of vegetation. Journal of Wind Engineering and Industrial
#  Aerodynamics 2017, 169, 168–176, /code{/link{doi:10.1016/j.jweia.2017.07.016.}}
# @examples
# /dontrun{
# P3D <- (1-(0.2/1.08))^(1/1.8)
# Lambda_V <- (1-P3D)*Lambda_V_1000m$pai
# }
# @rdname planar_area_index
# @export
planar_area_index <- function(lambda_building, lambda_vegetaion, P3D){
  # calculate the plane area index for the vegetation
  Lambda_V <- (1-P3D)*lambda_vegetaion
  # combine Lambda_building and lambda vegetation
  Lambda_p <- (Lambda_V+lambda_building)
  return(Lambda_p)
}


```

### frontal_area_index function 

```{r frontal_area_index function, include = FALSE }

# @title Calculate the frontal area index including buildings but also vegetaion
# as described by Kent et al. 2017
# @description The function used the frontal area index for vegetation and buildings
# (each evaluated spereatly by UMEP). The frontal area index for vegetation is
# further modified by the porosity.
# @param fai_v The frontal area index for vegetation.
# @param fai_b The fontral area index for buíldings.
# @param P3D The porosity (for leaf-on  P3D = 0.2, leaf-off P3D = 0.6 and intermediate P3D = 0.4)
# @param Cdb The drag coefficient  (Cdv) from Hagen and skidmore 1971
# single tree and one row deciduous windbreaks 0.5, vs. coniferous windbreaks 0.6-1.2)
# @return The function returns a new created frontal area index including buildings and
# vegetation.
# @examples
# /dontrun{
# frontal_area_index(fai_v = Lambda_V_1000m$fai, fai_b = Lambda_B_1000m$fai,
#                   Cdb = 1.2)
#                   }
# @seealso Kent, C. W.; Grimmond, S.; Gatey, D. Aerodynamic roughness parameters in
#  cities: Inclusion of vegetation. Journal of Wind Engineering and Industrial
#  Aerodynamics 2017, 169, 168–176, /code{/link{doi:10.1016/j.jweia.2017.07.016.}}
# @rdname frontal_area_index
# @export

frontal_area_index <- function(fai_v, fai_b, P3D,Cdb){
  Pv <- ((-1.251*(P3D^2)+0.489*(P3D)+0.803)/1*Cdb)
  lambda_output <- fai_b + (Pv*fai_v)
  return(lambda_output)
}


```

### Kanda_zd function 

```{r Kanda_zd function, include = FALSE }

# @title Calculates the z0 according to Kanda 2013 and Kent 2017
# @description The function estimates the zd (zero plane displacement) based on the approach
# described by Kanda 2013. Which is based on the method from MacDonald.
# @param pai The plane area index
# @param fai The frontal area index
# @param zH The average roughness elements height
# @param zstd The standard devitaion of the roughness elements height
# @param zHmax The maximum roughness element height
# @param Cd The drag coefficient  (Cdv) from Hagen and skidmore 1971
# single tree and one row deciduous windbreaks 0.5, vs. coniferous windbreaks 0.6-1.2), Default: 1.2
# @return The function returns the zero plane displacement zd as described in Kanda 2013,
# Kent 2017 and MacDonald 1998.
# @seealso Kanda, M.; Inagaki, A.; Miyamoto, T.; Gryschka, M.; Raasch, S.
# A New Aerodynamic Parametrization for Real Urban Surfaces. Boundary-Layer
# Meteorology 2013, 148, 357–377, /code{/link{doi:10.1007/s10546-013-9818-x.}}
# Kent, C. W.; Grimmond, S.; Gatey, D. Aerodynamic roughness parameters in
# cities: Inclusion of vegetation. Journal of Wind Engineering and Industrial
# Aerodynamics 2017, 169, 168–176, /code{/link{doi:10.1016/j.jweia.2017.07.016.}}
# Macdonald, R. W.; Griffiths, R. F.; Hall, D. J. An improved method for the
# estimation of surface roughness of obstacle arrays. Atmospheric Environment
# 1998, 32, 1857–1864, /code{/link{doi:10.1016/S1352-2310(97)00403-2.}}
# @examples
# /dontrun{
# # function to calculate the zd according to Kanda
# # Zd = (c0 * X^2 + ( (a0 * (Lp^b0) - c0) * X ) ) * Hmax
# # with a0 = 1.29, b0 = 0.36, c0 = -0.17 and X = ((omegaH + Haverage)/Hmax)
# Kanda_zd(pai = Lambda_B_1000m$pai,fai = Lambda_B_1000m$fai,
# zH = Lambda_B_1000m$zH, zstd = Lambda_B_1000m$zHstd,
# zHmax = Lambda_B_1000m$zHmax,Cd = 1.2)
# }
# @rdname Kanda_zd
# @export

Kanda_zd <- function(pai, fai, zH, zstd, zHmax, Cd = 1.2){
  # Basic params needed:
  Alph <- 4.43
  Beet <- 1.0
  k <- 0.4 # Karman Konstante
  # From Kanda
  Ao <- 1.29
  Bo <- 0.36
  Co <- -0.17

  # Run the Mcdonald method which is the basis for the Kanda method
  z_d_output <- (1+ (Alph^(-pai)) * (pai-1))*zH
  # Macexp <- 0.5*Beet*(Cd/k^2)*(1-z_d_output/zH)*fai
  # Calc the X
  X <- (zstd+zH)/zHmax
  # The fundamental MacDonald zd has to be adjusted by the Kanda method
  if(!is.na(X)){
    if(X >0 & X <= 1){
      z_d_output <- ((Co*(X^2))+((Ao*(pai^Bo)) -Co)*X)*zHmax
    }else{
      z_d_output <- (Ao*(pai^Bo))*zH
    }
  }else{
    z_d_output <- (Ao*(pai^Bo))*zH
  }

  return(z_d_output)
}


```

### Kanda_z0 function 

```{r Kanda_z0 function, include = FALSE }

# @title Calculates the z0 according to Kanda 2013 and Kent 2017
# @description The function estimates the z0 (roughness length) based on the approach
# described by Kanda 2013. Which is based on the method from MacDonald.
# @param fai The frontal area index.
# @param pai The plane area index
# @param zstd The standard deviation of the roughness elements height
# @param zH The average roughness elements height
# @param zHmax The maximum roughness elements height
# @param Cd The drag coefficient  (Cdv) from Hagen and skidmore 1971
# single tree and one row deciduous windbreaks 0.5, vs. coniferous windbreaks 0.6-1.2), Default: 1.2
# @return The function returns a z0 including vegetation and buildings described in the
# method from Kanda 2013, Kent 2017 and MacDonald 1998
# @seealso Kanda, M.; Inagaki, A.; Miyamoto, T.; Gryschka, M.; Raasch, S.
# A New Aerodynamic Parametrization for Real Urban Surfaces. Boundary-Layer
# Meteorology 2013, 148, 357–377, /code{/link{doi:10.1007/s10546-013-9818-x.}}
# Kent, C. W.; Grimmond, S.; Gatey, D. Aerodynamic roughness parameters in
# cities: Inclusion of vegetation. Journal of Wind Engineering and Industrial
# Aerodynamics 2017, 169, 168–176, /code{/link{doi:10.1016/j.jweia.2017.07.016.}}
# Macdonald, R. W.; Griffiths, R. F.; Hall, D. J. An improved method for the
# estimation of surface roughness of obstacle arrays. Atmospheric Environment
# 1998, 32, 1857–1864, /code{/link{doi:10.1016/S1352-2310(97)00403-2.}}
# @examples
# /dontrun{
# # Zd = (c0 * X^2 + ( (a0 * (Lp^b0) - c0) * X ) ) * Hmax
# # with a0 = 1.29, b0 = 0.36, c0 = -0.17 and X = ((omegaH + Haverage)/Hmax)
# # Z0 = ((b1*Y^2) + (c1*Y) + a1)MacZ0
# # Y <- ((Lambda_summer*zH_std)/zH)
# # Cdb <- 1.2 # Hall et al 1996 (from Kent et al 2017a)
# Kanda_z0(fai = Lambda_B_1000m$fai,pai = Lambda_B_1000m$pai,
#          zstd = Lambda_B_1000m$zHstd, zH = Lambda_B_1000m$zH,
#          zHmax = Lambda_B_1000m$zHmax,Cd = 1.2)
#          }
# @rdname Kanda_z0
# @export

Kanda_z0 <- function(fai, pai, zstd, zH, zHmax, Cd = 1.2){
  # Basic params needed:
  Alph <- 4.43
  Beet <- 1.0
  k <- 0.4 # Karman Konstante

  A1 <- 0.71
  B1 <- 20.21
  C1 <- -0.77
  # THe basic zd from McDonald is needed for the calculation of the z0Mac
  z_d_output <- (1+ (Alph^(-pai)) * (pai-1))*zH
  z0Mac <- (zH*(1-z_d_output/zH)*exp(-(0.5*(Cd/k^2)*(1-(z_d_output/zH))*fai)^-0.5))
  # This z0Mac is than adjusted by the Kanda method
  Y <- (pai*zstd/zH)
  if(!is.na(Y)){
    if(Y >= 0){
      z_0_output <- ((B1*(Y^2))+(C1*Y)+A1)*z0Mac
    }else{
      z_0_output <- A1*z0Mac
    }
  }else{
    z_0_output <- A1*z0Mac
  }


  return(z_0_output)
}


```

### RUN_Kanda_calc

```{r Correct UMEP data for seasonality }

# from the UMEP I will get one zd and z0 determined for the buildings as obstrucles
# and anotherone for the vegetation. The combination of both will than be used for 
# the determination of the "REAL" zd and z0 sperearted into a leaf-off and -on seasons.

# (Cdv = drag coefficient for vegetation: from Hagen and skidmore 1971
  # single tree and one row deciduous windbreaks 0.5, vs. coniferous windbreaks 
  # 0.6-1.2)
# Kent 2017b: uses for leaf-on  P3D = 0.2 and leaf-off P3D = 0.6 (and 0.4 for intermediate)
    # do the Lp_summer and winter also for each degree 
# summer
    pai_deg_summer <- planar_area_index(lambda_building = Lambda_B_1deg$pai, 
                             lambda_vegetaion = Lambda_V_1deg$pai,P3D = 0.2)
# winter
    pai_deg_winter <- planar_area_index(lambda_building = Lambda_B_1deg$pai, 
                                 lambda_vegetaion = Lambda_V_1deg$pai,P3D = 0.6)
# intermediate
    pai_deg_intermediate <- planar_area_index(lambda_building = Lambda_B_1deg$pai, 
                                        lambda_vegetaion = Lambda_V_1deg$pai,P3D = 0.4)
    
# and easy calculation of the frontal area fraiton

# and for each degree
fai_deg_summer <- frontal_area_index(fai_v = Lambda_V_1deg$fai,
                                     fai_b = Lambda_B_1deg$fai, 
                                     P3D = 0.2, Cdb = 1.2)

fai_deg_winter <- frontal_area_index(fai_v = Lambda_V_1deg$fai,
                                 fai_b = Lambda_B_1deg$fai, 
                                 P3D = 0.6, Cdb = 1.2)

fai_deg_intermediate <- frontal_area_index(fai_v = Lambda_V_1deg$fai,
                                     fai_b = Lambda_B_1deg$fai, 
                                     P3D = 0.4, Cdb = 1.2)

# Calcualte the zd and z0 from Lp_summer, and winter from Kanda:
  # This should be the same as in Lambda_B_1000m$zd
    # Zd_compare = Kanda_zd(pai = Lambda_B_1000m$pai,fai = Lambda_B_1000m$fai,
    #                       zstd = Lambda_B_1000m$zHstd,zH = Lambda_B_1000m$zH,
    #                       zHmax = Lambda_B_1000m$zHmax,Cd = 1.2)

# Zd summer lies inbetween zd_building and zd_vegetation (as it should be!)

# and now for each degree
Zd_deg_summer <- c()
Zd_deg_winter <- c()
Zd_deg_intermediate <- c()
for(i in 1:360){
  Zd_deg_summer[i] <- Kanda_zd(pai = pai_deg_summer[i], fai = fai_deg_summer[i],
                            zH = Lambda_comb_1deg$zH[i], zstd = Lambda_comb_1deg$zHstd[i],
                            zHmax = Lambda_comb_1deg$zHmax[i])
    Zd_deg_winter[i] <- Kanda_zd(pai = pai_deg_winter[i], fai = fai_deg_winter[i],
                            zH = Lambda_comb_1deg$zH[i], zstd = Lambda_comb_1deg$zHstd[i],
                            zHmax = Lambda_comb_1deg$zHmax[i])
    Zd_deg_intermediate[i] <- Kanda_zd(pai = pai_deg_intermediate[i], 
                                       fai = fai_deg_intermediate[i],
                                 zH = Lambda_comb_1deg$zH[i], 
                                 zstd = Lambda_comb_1deg$zHstd[i],
                                 zHmax = Lambda_comb_1deg$zHmax[i])
}
# difference plot (influence of season on roughness)
# plot(Zd_deg_summer-Zd_deg_winter)

# And the same for the z0
# and now for each degree
Z0_deg_summer <- c()
Z0_deg_winter <- c()
Z0_deg_intermediate <- c()
for(i in 1:360){
  Z0_deg_summer[i] <- Kanda_z0(fai = fai_deg_summer[i], pai = pai_deg_summer[i], 
                               zstd = Lambda_comb_1deg$zHstd[i], zH = Lambda_comb_1deg$zH[i],
                               zHmax = Lambda_comb_1deg$zHmax[i])
  Z0_deg_winter[i] <- Kanda_z0(fai = fai_deg_winter[i], pai = pai_deg_winter[i], 
                               zstd = Lambda_comb_1deg$zHstd[i], zH = Lambda_comb_1deg$zH[i],
                               zHmax = Lambda_comb_1deg$zHmax[i])
  Z0_deg_intermediate[i] <- Kanda_z0(fai = fai_deg_intermediate[i], pai = pai_deg_intermediate[i], 
                                     zstd = Lambda_comb_1deg$zHstd[i], zH = Lambda_comb_1deg$zH[i],
                                     zHmax = Lambda_comb_1deg$zHmax[i])
}
# plot(Z0_deg_summer-Z0_deg_winter)

# make data table to export
# for summer
  summer_1deg <- data.table::data.table(cbind(c(0:359),pai_deg_summer, fai_deg_summer, 
                                  Zd_deg_summer, Z0_deg_summer))
  names(summer_1deg)[1] <- "Wd"
# and for winter
  winter_1deg <- data.table::data.table(cbind(c(0:359),pai_deg_winter, fai_deg_winter, 
                                  Zd_deg_winter, Z0_deg_winter))
  names(winter_1deg)[1] <- "Wd"
# and for intermediate
  intermediate_1deg <- data.table::data.table(cbind(c(0:359),pai_deg_intermediate, fai_deg_intermediate, 
                                  Zd_deg_intermediate, Z0_deg_intermediate))
  names(intermediate_1deg)[1] <- "Wd"


```

# Running mean calculation 

**Step 3)** Average the one-degree zd values over a running mean including the zd values from wind direction i±5 degree moving forward by one degree.

Justus: "I created a lookup table for mean values for each possible wind directions (1-360). The needed `running_mean` function was until now not included in the UCON Package but i just uploaded it. The other functions `order_zd_z0` and `average_z_by_wd` are used afterwards to arrange the numbers from the lookup table to the according wd.. Depending on the Method (Kotthaus wd+v_sd or wd+-5°) different lookup tables should be used!"  

Why use a running mean? It smooths in a 10-degree slice: 1-degree is too narrow to be representative. Source? Kotthaus and Grimmond 2014 used a similar (though not the same) ‘running mean’ method. The 10-degree slice was an idea from Fred. 

## Running mean function 

```{r running_mean function, include = FALSE }

# @title Creates a running mean for data e.g. roughness parameter
# @description The function takes data that should be in a radial degree order
# and averages the data for each degree ± 5 degree.
# @param data This should be a data.table with 360 rows each including the information
# for the 1 degree.
# @return The function returns a data.table with again 360 rows but each including now
# the average at point i±5.
# @examples
# mean(unlist(summer_1deg[c((355:360),(1:5)),..a]))
# summer_1deg_running <- running_mean(data = summer_1deg)
# summer_1deg_running$Z0_deg_summer[360]
# @rdname running_mean
# @export

running_mean <- function(data){
  if(nrow(data) != 360){
    stop("The data should have 360 rows, one for each degree!")
  }
  output_mean <- data.table::data.table(matrix(as.numeric(), ncol = ncol(data), nrow = 360))
  names(output_mean) <- names(data)
  for(i in 1:360){
    for(a in 2:ncol(data)){
      if(i <5){
        output_mean[i,a] <- mean(unlist(data[c(((360+i-5):360),(0:(i+5))),..a]))
      }else if(i > 355 & i < 360){
        output_mean[i,a] <- mean(unlist(data[c(((i-5):i),(i+1):360,(1:(5+i-360))),..a]))
      }else if(i == 360){
        output_mean[i,a] <- mean(unlist(data[c((355:360),(1:5)),..a]))
      }else{
        output_mean[i,a] <- mean(unlist(data[c((i-5):(i+5)),..a]))
      }
    }
  }
  output_mean[,1] <- c(1:360)
  return(output_mean)
}



```

## Apply running_mean 

```{r Apply running_mean }

summer_1deg_running = running_mean(data = summer_1deg)
head(summer_1deg_running)

# plot original data and running_mean (summer zd)
ggplot()+
  geom_line(data = summer_1deg, aes(x = Wd, y = Zd_deg_summer), color = "blue")+
  geom_line(data = summer_1deg_running, aes(x = Wd, y = Zd_deg_summer), color = "red")

winter_1_deg_running = running_mean(data = winter_1deg)

intermediate_1_deg_running = running_mean(data = intermediate_1deg)

```

# Eddy flux data 

Load the eddy flux data. 

```{r Load eddy flux data }

TUCC_raw <- read.delim("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/03_RAW_DATA/reeddyfluxdata2020/UCO_TUCC_for_UWI_2014-07-03_2021-05-31.txt", header = TRUE, sep=",")

glimpse(TUCC_raw)

```

## Simplify column names 

The column names are changed (shorter).

```{r Change column names }

colnames(TUCC_raw)

names(TUCC_raw)[names(TUCC_raw) == "timestamp..UTC."] <- "timestamp"

names(TUCC_raw)[names(TUCC_raw) == "upward.sensible.heat.flux.in.air..W.m.2...H...56.17m."] <- "H"

names(TUCC_raw)[names(TUCC_raw) == "quality.flag.upward.sensible.heat.flux.in.air..cat...qc_H...56.17m."] <- "qc_H"

names(TUCC_raw)[names(TUCC_raw) == "upward.latent.heat.flux.in.air..W.m.2...LE...56.17m."] <- "LE"

names(TUCC_raw)[names(TUCC_raw) == "quality.flag.upward.latent.heat.flux.in.air..cat...qc_LE...56.17m."] <- "qc_LE"

names(TUCC_raw)[names(TUCC_raw) == "H2O.signal.strength.infrared.gasanalyser...1...H2O_sig_strgth_mean...56.17m."] <- "H2O_sig_strgth_mean"

names(TUCC_raw)[names(TUCC_raw) == "evapotranspiration.rate..mm.h.1...ET...56.17m."] <- "ET"

names(TUCC_raw)[names(TUCC_raw) == "upward.mole.flux.of.carbon.dioxide.in.air..1e.6.mol.s.1.m.2...co2_flux...56.17m."] <- "co2_flux"

names(TUCC_raw)[names(TUCC_raw) == "CO2.signal.strength.infrared.gasanalyser..1...CO2_sig_strgth_mean...56.17m."] <- "CO2_sig_strgth_mean"

names(TUCC_raw)[names(TUCC_raw) == "wind.speed..m.s.1...wind_speed...56.17m."] <- "wind_speed"

names(TUCC_raw)[names(TUCC_raw) == "wind.from.direction..degree...wind_dir...56.17m."] <- "wind_dir"

names(TUCC_raw)[names(TUCC_raw) == "friction.velocity..m.s.1...u....56.17m."] <- "friction_velocity"

names(TUCC_raw)[names(TUCC_raw) == "obukhov.length..m...L...56.17m."] <- "obukhov_length"

names(TUCC_raw)[names(TUCC_raw) == "monin.obukhov.stability.parameter..1....z.d..L...56.17m."] <- "monin_obukhov_stability_parameter"

names(TUCC_raw)[names(TUCC_raw) == "bowen.ratio..1...bowen_ratio...56.17m."] <- "bowen_ratio"

names(TUCC_raw)[names(TUCC_raw) == "northward.wind..m2.s.2...v_var...56.17m."] <- "northward_wind"

names(TUCC_raw)[names(TUCC_raw) == "w.wind.component..m.s.1...w_unrot...56.17m."] <- "wind_unrot"

names(TUCC_raw)[names(TUCC_raw) == "air.temperature..degree_C...T_40m_Avg...56.17m."] <- "air_temperature"

names(TUCC_raw)[names(TUCC_raw) == "relative.humidity......RH_40m_Avg...56.17m."] <- "RH"

names(TUCC_raw)[names(TUCC_raw) == "air.pressure..hPa...air_pressure...56.17m."] <- "air_pressure"

names(TUCC_raw)[names(TUCC_raw) == "liquid.precipitation.rate..mm.h.1...NS_Int_fluid...1.00m."] <- "precipitation"

names(TUCC_raw)[names(TUCC_raw) == "downwelling.shortwave.flux.in.air..W.m.2...SPN_TOTAL_Avg...56.17m."] <- "SPN_TOTAL_Avg"

names(TUCC_raw)[names(TUCC_raw) == "diffuse.downwelling.shortwave.flux.in.air..W.m.2...SPN_DIFFUSE_Avg...56.17m."] <- "diffuse_radiation"

names(TUCC_raw)[names(TUCC_raw) == "downwelling.shortwave.flux.in.air..W.m.2...Rs_downwell_Avg...56.17m."] <- "radiation.shortwave.in"

names(TUCC_raw)[names(TUCC_raw) == "downwelling.longwave.flux.in.air..W.m.2...Rl_downwell_Avg...56.17m."] <- "radiation.longwave.in"

names(TUCC_raw)[names(TUCC_raw) == "upwelling.shortwave.flux.in.air..W.m.2...Rs_upwell_Avg...56.17m."] <- "radiation.shortwave.out"

names(TUCC_raw)[names(TUCC_raw) == "upwelling.longwave.flux.in.air..W.m.2...Rl_upwell_Avg...56.17m."] <- "radiation.longwave.out"

names(TUCC_raw)[names(TUCC_raw) == "surface.net.downward.radiative.flux..W.m.2."] <- "net.downward.radiative"

```

## Filter eddy flux data for quality 

I am leaving the whole dataset now to run the entire footprint analyses on it! 

```{r Filter for quality }

# Convert NaN to NA 
TUCC_raw[ TUCC_raw == "NaN"] = NA # Watch out - the new EC data uses both NaN and NAN
TUCC_raw[ TUCC_raw == "NAN"] = NA

# COnvert non-date columns to numeric 
# https://stackoverflow.com/questions/3796266/change-the-class-from-factor-to-numeric-of-many-columns-in-a-data-frame
TUCC_raw[,2:30] = apply(TUCC_raw[,2:30], 2, function(x) as.numeric(as.character(x)))

#TUCC_raw = TUCC_raw[!(TUCC_raw$qc_LE == 2), ]
# I will do data filtering later. 

```

## Reduce to 2018-2021

Years 2018-2021 for my 3rd paper 

```{r reduce years}

TUCC_raw$timestamp = as.POSIXct(TUCC_raw$timestamp, tz = "GMT")

TUCC_raw$Year = lubridate::year(TUCC_raw$timestamp)

TUCC_18_21 = subset(TUCC_raw, Year == 2018 | Year == 2019 |
                      Year == 2020 | Year == 2021 )

```


## Arrange zd from lookup table according to wd 

"I created a lookup table for mean values for each possible wind directions (1-360). The needed running_mean function was until now not included in the UCON Package but i just uploaded it. The other functions order_zd_z0 and average_z_by_wd are used afterwards to arrange the numbers from the lookup table to the according wd.. Depending on the Method (Kotthaus wd+v_sd or wd+-5°) different lookup tables should be used!"

### average_z_by_wd function 

```{r average_z_by_wd function, include = FALSE }

# @title Selects zd and z0 accoring to the prevailing wind direction and uses as zd and z0
# @description The basis is a running mean lookup-table or as described in the method
# by Kothaus and Grimmond 2014 where they average zd and z0 for the WD degree ± sd_v.
# @param wd Wind direction
# @param v_sd If the Kotthaus Method is used sd_v is needed otherwise not,thus Default: NULL
# @param z_data The z_data should be a data.table with the zd and z0 values
# (as data.table("deg"= c(1:360), "zd"= Zd_winter, "z0" = Z0_winter)).
# @param deg_column The name of the columne in z_data where the degree is.
# @param Kotthaus Logical if the Kotthaus mehtod is used the default is FALSE!
# @return The function returns a vector with two numbers. The first one ist the
# averaged zd and the second one the z0 value.
# @seealso Kotthaus, S.; Grimmond, C. S. B. Energy exchange in a dense urban
# environment – Part II: Impact of spatial heterogeneity of the surface. Urban
# Climate 2014, 10, Part 2, 281–307, /code{/link{doi:10.1016/j.uclim.2013.10.001.}}
# /code{/link{oder_zd_z0}}
# @examples
# # explain the function
# /dontrun{
# min <- round(all_day_footprint$wind_dir[120], 0)
# max <- round(all_day_footprint$wind_dir[120],0)
# mean(z_summer$`zm-zd`[which(z_summer$deg <= max &
#                            z_summer$deg >= min)])
# # test the function
# average_z_by_wd(wd = all_day_footprint$wind_dir[120],
#                 z_data = z_summer,deg_column = "deg", z_column= "`zm-zd`")
#                 }
# @rdname average_z_by_wd
# @export
#
average_z_by_wd <- function(wd, v_sd= NULL, z_data, deg_column, Kotthaus = F){
  # create the winddirection range where to average over
  if(Kotthaus == F){
    # in case that the wd+v_sd is above 360 degree
    if(round(wd,0) >360){
      max_wd <- round(round((wd)-360,0),0)
      min_wd <- round(wd,0)
    }else if(round(wd,0) < 1){ # and in case that the wd+v_sd is below 1 deg
      min_wd <- round(round((wd)+360,0))
      max_wd <- round(wd,0)
    }else{
      min_wd <- round(wd,0)
      max_wd <- round(wd,0)
    }
  }else{
    # in case that the wd+v_sd is above 360 degree
    if(round(wd+v_sd,0) >360){
      max_wd <- round(round((wd+v_sd)-360,0),0)
      min_wd <- round(wd-v_sd,0)
    }else if(round(wd-v_sd,0) < 1){ # and in case that the wd+v_sd is below 1 deg
      min_wd <- round(round((wd-v_sd)+360,0))
      max_wd <- round(wd+v_sd,0)
    }else{
      min_wd <- round(wd-v_sd,0)
      max_wd <- round(wd+v_sd,0)
    }
  }

  # select the z_data according to max_wd and min_wd
  output_1 <- mean(z_data$zd[which(z_data$deg <= max_wd &
                              z_data$deg >= min_wd)])
  output_2 <- mean(z_data$z0[which(z_data$deg <= max_wd &
                                          z_data$deg >= min_wd)])
  return(c(output_1,output_2))
}


```

### order_zd_z0 function

```{r order_zd_z0 function, include = FALSE }

# @title Orders the zd and z0 data by the wd for a given dataset
# @description Two options are possible how to average zd and z0. First and default
# is to take zd and z0 from a lookup-table including the running mean for each degree
# that is a mean for the degree ±5°. The other option is the method described
# by Kothaus and Grimmond 2014 where they average zd and z0 for the WD degree ± sd_v.
# @param Zd_winter Zero plane displacement (zd) data for winter
# @param Z0_winter Roughness length (z0) data for winter
# @param Zd_intermediate Zero plane displacement (zd) data for intermediat
# @param Z0_intermediate Roughness length (z0) data for intermediate
# @param Zd_summer Zero plane displacement (zd) data for summer
# @param Z0_summer Roughness length (z0) data for summer
# @param Timestamp A simple Timestamp to indentify the seson.
# @param Kotthaus Decide which method should be used. The default is the running mean, Default: F
# @param wind_direction The wind direction data.
# @param wind_v In case the Kotthaus method is applied, sd_v is needed and should be applied
# here. Otherwise just, Default: NULL
# @return The functoin returns a vector of same length as all input data including the
# zd and z0 depending on the selected method.
# @seealso Kotthaus, S.; Grimmond, C. S. B. Energy exchange in a dense urban
# environment – Part II: Impact of spatial heterogeneity of the surface. Urban
# Climate 2014, 10, Part 2, 281–307, /code{/link{doi:10.1016/j.uclim.2013.10.001.}}
# /code{/link{average_z_by_wd}}
# @examples
# /dontrun{
# rralle <- order_zd_z0(Zd_winter = winter_1deg_500$Zd_deg_winter_500,
#              Z0_winter = winter_1deg_500$Z0_deg_winter_500,
#              Zd_intermediate = intermediate_1deg_500$Zd_deg_intermediate_500,
#              Z0_intermediate = intermediate_1deg_500$Z0_deg_intermediate_500,
#              Zd_summer = summer_1deg_500$Zd_deg_summer_500,
#              Z0_summer = summer_1deg_500$Z0_deg_summer_500,
#              Timestamp = all_day_footprint$TIMESTAMP,
#              wind_direction = all_day_footprint$wind_dir)
#              }
# @rdname order_zd_z0
# @export
order_zd_z0 <- function(Zd_winter, Z0_winter, Zd_intermediate, Z0_intermediate,
         Zd_summer, Z0_summer, Timestamp, Kotthaus = F,
         wind_direction, wind_v = NULL){
  # check if the length of wind_direction and Timestamp are equal
  if(length(Timestamp) != length(wind_direction) ){
    stop("The length of Timestamp and wind_direction differ!")
  }
  # generate look-up data.table
  z_winter <- data.table::data.table("deg"= c(1:360),"zd"= Zd_winter,
                         "z0" = Z0_winter)
  z_intermediate <- data.table::data.table("deg"= c(1:360),"zd"= Zd_intermediate,
                               "z0" = Z0_intermediate)
  z_summer <- data.table::data.table("deg"= c(1:360),"zd"= Zd_summer,
                         "z0" = Z0_summer)
  # create an empty output data.table
  z_result <- data.table::data.table(TIMESTAMPE = Timestamp, zd= rep(as.numeric(NA),length(wind_direction)),
                         z0= rep(as.numeric(NA),length(wind_direction)))

  # in case that the average is made with the running mean
  if(Kotthaus == F){
    for(i in 1:nrow(z_result)){
      if(is.na(wind_direction[i])){
        z_result$zd[i] <- NA
        z_result$z0[i] <- NA
      }else{
        if(substr(Timestamp[i],6,7) %in% c("12","01","02")){  # THIS IS WINTER
            res <- average_z_by_wd(wd = wind_direction[i],
                                   z_data = z_winter, deg_column = "deg")
          # res includes two different data 1 = zd and 2 = z0
          z_result$zd[i] <- res[1]
          z_result$z0[i] <- res[2]
        }else if(substr(Timestamp[i],6,7) %in% c("06","07","08")){ # THIS IS SUMMER
            res <- average_z_by_wd(wd = wind_direction[i],
                                   z_data = z_summer, deg_column = "deg")
          z_result$zd[i] <- res[1]
          z_result$z0[i] <- res[2]
        }else{
            res <- average_z_by_wd(wd = wind_direction[i],
                                   z_data = z_intermediate, deg_column = "deg")
          z_result$zd[i] <- res[1]
          z_result$z0[i] <- res[2]
        }
      }
    }
  }else{
    for(i in 1:nrow(z_result)){
      if(is.na(wind_direction[i])){
        z_result$zd[i] <- NA
        z_result$z0[i] <- NA
      }else{
      if(substr(Timestamp[i],6,7) %in% c("12","01","02")){  # THIS IS WINTER
          res <- average_z_by_wd(wd = wind_direction[i],v_sd = wind_v[i],
                                 z_data = z_winter, deg_column = "deg", Kotthaus = T)
        # res includes two different data 1 = zd and 2 = z0
        z_result$zd[i] <- res[1]
        z_result$z0[i] <- res[2]
      }else if(substr(Timestamp[i],6,7) %in% c("06","07","08")){ # THIS IS SUMMER
          res <- average_z_by_wd(wd = wind_direction[i],v_sd = wind_v[i],
                                 z_data = z_summer, deg_column = "deg", Kotthaus = T)
        z_result$zd[i] <- res[1]
        z_result$z0[i] <- res[2]
      }else{
          res <- average_z_by_wd(wd = wind_direction[i],v_sd = wind_v[i],
                                 z_data = z_intermediate, deg_column = "deg", Kotthaus = T)
        z_result$zd[i] <- res[1]
        z_result$z0[i] <- res[2]
      }
    }
    }
  }
  return(z_result)
}


```

### Apply order_zd_z0 

```{r apply order_zd_z0 }

# Use the running means generated earlier 
zd_z0_TUCC = order_zd_z0(Zd_winter = winter_1_deg_running$Zd_deg_winter,
                         Z0_winter = winter_1_deg_running$Z0_deg_winter,
                         Zd_intermediate = intermediate_1_deg_running$Zd_deg_intermediate,
                         Z0_intermediate = intermediate_1_deg_running$Z0_deg_intermediate,
                         Zd_summer = summer_1deg_running$Zd_deg_summer,
                         Z0_summer = summer_1deg_running$Z0_deg_summer,
                         Timestamp = TUCC_18_21$timestamp,
                         wind_direction = TUCC_18_21$wind_dir
)

glimpse(zd_z0_TUCC)
# You get a dataframe with 3 columns: TIMESTAMP, zd, and z0, which corresponds to the eddy flux dataframe. 
# You can easily merge this result to the original eddy flux dataframe. 

```

### Combining data 

Now, we merge the zd/z0 with the original eddy flux dataset. 

```{r combining data }

names(zd_z0_TUCC)[names(zd_z0_TUCC) == "TIMESTAMPE"] <- "timestamp"

TUCC_final = dplyr::full_join(TUCC_18_21, zd_z0_TUCC)

glimpse(TUCC_final)

```

Let's save the data. 

```{r save data }

#save(TUCC_final, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211014_zd_firstguess/TUCC/20211014_TUCC_final.RData")

#write.csv(TUCC_final, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211014_zd_firstguess/TUCC/20211014_TUCC_final.csv", row.names = FALSE)

```

