---
title: "Performance metrics (TUCC 2020)"
author: "Stenka Vulova"
date: "22 Nov. 2021"
output:
  pdf_document: default
  html_document: default
---

I will calculate the performance metrics for TUCC 2020.  
I am comparing the weighted averages of footprints extracted from the area around the towers to the EC data.  

The EC data went through quality control. 


# Libraries 

```{r libraries, include  = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(FSA) # Dunns test
library(RColorBrewer)
library(reshape2)
library(scales)
library(hydroGOF) # performance metrics

library(janitor)

# machine learning libraries 
library(caret)
#library(CAST)
library(doSNOW) # do both_readying in parallel
library(doParallel) 
#library(gbm)
library(AppliedPredictiveModeling)
library(randomForest)
#library(nnet) # avNNet

# Benchmarking 
library(tictoc)

```

# Load EC data 

```{r load EC dataset }

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/04_PROCESSED_DATA/20211026_QC/TUCC/06_add_NDVI/TUCC_w_NDVI.RData")

glimpse(TUCC_w_NDVI)

EC_ET = TUCC_w_NDVI %>% 
  dplyr::select(timestamp, ET)

# Subset to 2020 
EC_ET$Year = lubridate::year(EC_ET$timestamp)

EC_ET_20 = subset(EC_ET, Year == 2020)

summary(EC_ET_20)
nrow(EC_ET_20) # 7776

```

# Load predicted ET 

It was extracted with footprints from predicted ET maps. 
The ET predictions (maps) were generated from a model trained in 2020. 

```{r load predicted ET }

load("/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/05_RESULTS/20211117_ETmaps_ftp/2020/TUCC/FP_TUCC_ET_20_df.RData")

summary(FP_TUCC_ET_20_df)

FP_ET = FP_TUCC_ET_20_df %>% 
  dplyr::select(timestamp, ET_FP)

# Merge the two 
preddf = dplyr::left_join(EC_ET_20, FP_ET)

glimpse(preddf)
summary(preddf)

```

# Metrics

```{r metrics }

caret::RMSE(preddf$ET, preddf$ET_FP) # 0.02107674

caret::R2(preddf$ET, preddf$ET_FP) # 0.4851754

```

## Metrics table 

```{r metrics table }

metrics_Test20_TUCC_RF = 
  data.frame(Tower_test = "TUCC",
              Year_test = "20",
             
             Model = "RF", 
             
             RMSE = hydroGOF::rmse(preddf$ET_FP, preddf$ET),
             
             ME = hydroGOF::me(sim = preddf$ET_FP, obs = preddf$ET),
             
             pbias = hydroGOF::pbias(sim = preddf$ET_FP, obs = preddf$ET),
             
             R2 = caret::R2(preddf$ET_FP, preddf$ET),
             
             NSE = hydroGOF::NSE(sim = preddf$ET_FP, obs = preddf$ET),

             MAE = hydroGOF::mae(sim = preddf$ET_FP, obs = preddf$ET), 

             NRMSE = hydroGOF::nrmse(sim = preddf$ET_FP, obs = preddf$ET, norm = "maxmin") )

metrics_Test20_TUCC_RF

# save

#write.csv(metrics_Test20_TUCC_RF, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/05_RESULTS/20211118_metrics_RF/Test2020/TestTUCC/metrics_Test20_TUCC_RF.csv", row.names = FALSE)

#save(metrics_Test20_TUCC_RF, file = "/Users/stenkavulova/Dropbox/Stenka_UWI/Topic_3/05_RESULTS/20211118_metrics_RF/Test2020/TestTUCC/metrics_Test20_TUCC_RF.RData")

```

